Is statistical significance a thing?

### 2018-02-16 doc.tex
Update doc.tex

\textbf{Abstract}: Decades of misuse and misinterpretation of Null Hypothesis Significance Testing (NHST) has led to recent 
calls for abandoning its use. We argue that statistical hypothesis tests are useful, but are not, and should not be 
referred to, as tests of significance. We propose that a sensible method to reduce misconceptions about p~values 
is to refer to the results of NHST (i.e. $p < or > 0.05$) using ``statistical clarity''. A test of the null hypothesis tells 
us whether or not we can \emph{clearly} see our effect (e.g. whether a correlation is positive or negative, or if a response 
in the treatment group is greater than the response in the control group). Specifically, we argue the use of ``statistical
clarity'' will help researchers avoid accepting the null hypothesis and retrieve the English definition for ``significance'' 
to shift focus back to substantive significance.

@@

For nearly a century statisticians and scientists have described the shortcomings of Null Hypothesis Significance Testing
(NHST) \citep[see][]{Cohen1994}. Research articles and books written in the past two decades proposing a deemphasis of the 
p~value in favor of effect sizes and confidence intervals have been cited thousands of times \citep{Cohen1994, Goodman1999, 
Wilkinson1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}. Yet, both a reliance on NHST and the misinterpretation of 
results obtained from NHST remain high. In a 2017 meta-analysis of 356 studies published in \emph{European Sociological 
Research}, \citet{Bernardietal.2017} found that 72\% of studies contained an ambiguous use of the term ``significant'', 
49\% interpreted non-significant effects as zero effects, and 44 \% failed to report an effect size in a comprehensible 
way. The misuse and misinterpretation of NHST is so pervasive that there has been a recent push for drastically reducing its 
use \citep{SzucsandIoannidis2017} or abandoning its use altogether \citep{McShaneetal.2017}. Others have called for a more
strenuous significance threshold (e.g. $p < 0.005$: \citealt{Benjaminetal.2018}); however, this change is unlikely to reduce
the misuse and misinterpretation of p~values, or the publication bias imposed by any p~value threshold 
\citep{Ridleyetal.2007}

Alternatively, we suggest a change in the language used in NHST. Results from NHST (i.e. $p < or > 0.05$) should be 
described using language that directly corresponds to what a test of the null hypothesis actually tells us: whether or not 
we can \emph{clearly} see our effect. For example, given $H_{0} = 0$ (e.g. the difference in our treatments is 0) what a 
$p < 0.05$ tells us is whether or not we can clearly see the sign of the effect. Using the language of ``statistical 
clarity'' would help researchers escape the trap of accepting the null hypothesis and reduce multiple types of
misinterpretation associated with accepting the null hypothesis, while allowing for it to be retained as a method of 
inference (e.g. see \citealt{Abelson1997} for arguments for NHST). For example,equating $p < 0.05$ with a zero effect would 
be much harder to make once a researcher has written the following sentence: ``The difference between our control and 
treatment group was not statistically clear ($p = 0.24$). Additionally, the use of $p > 0.05$ as evidence for something 
(accepting the null hypothesis) would be more difficult if the following language was used: ``We were unable to clearly 
determine how our control and sham treatments differed''. Additionally, once ``significant'' regained its common English 
definition, conflating $p < 0.05$ with substantive significance would be more difficult.

\begin{itemize}
  \item Other topics to mention
  	\subitem difficulty in interpreting confidence intervals?
    \subitem usefulness of p --- use in splines: non-ridiculous standard to determine if we can see a pattern
    \subitem a p~value is not a property of reality (a biological ``thing''), but a property of a data set and assumptions...
\end{itemize}

@@

\mk{Recount the way a high p is obtained?}

\subsection*{$\mathbf{p > 0.05 \neq}$ zero effect}

In the vast majority of scientific fields a truly zero, ``null effect'' does not exist \citep{Meehl1990, Tukey1991,
Cohen1994}. Yet, references to a null effect can still be found in high impact journals \citep[e.g.][]{Feldonetal.2017}. 
To emphasize that effects are not truly zero, descriptions of statistical results using the language of clarity should
begin with reference to the effect. For example, ``The difference between the control and treatment group was not 
statistically clear.''

\begin{itemize}
  \item Equating $p > 0.05$ with no difference: \citep{Ortegoetal.2007, Singhetal.2017}
\end{itemize}

\mk{Probably do not need this full description here}

\begin{itemize}
  \item ``...establish the \emph{a priori} selection of an effect level that is considered unacceptably toxic:''
  \citep{Dentonetal.2011}
\end{itemize}

\subsection*{Using larger p~values as more evidence for something}

Another common mistake is using a larger p~value as stronger evidence for accepting the null hypothesis.  For example, 
\mk{Need a good quote here. An ok placeholder:} ``...yielded nonsignificant differences on measures of those skills with P 
values of at least 0.2, and most in the $0.7 \leq P \leq 0.9$ range.'' \citep{Feldonetal.2017}. A larger p value actually
corresponds to less clarity and thus the possibility of an undetected larger effect.

\mk{Point about focal vs non-focal hypotheses. Often accept the null for non-focal (diagnostics). While this is not
appropriate language it seems less bad...}

\begin{itemize}
  \item Setting out to accept the null: \citep{Karulinetal.2015}
\end{itemize}

\mk{Again, maybe this is too much}

\subsection*{The difference between clear and not clear is not clear}

In 2006 Gelman and Stern wrote that the difference between statistically significant and statistically non-significant is
not itself significant. Rewritten using ``clarity'' this statement indicates that if a single result is not statistically 
clear, any comparison made between that unclear result and a clear result will not itself be a clear comparison. An example
taken from the literature is the following:

For example, if a treatment induces a statistically clear increase in growth rate in males, but has an unclear effect in 
females, the relationship between the treatment on males and females is not clear.

\begin{itemize}
  \item With the \emph{correlation coefficient} for males and females (but significance for only one sex---``substrate 
  choice correlated significantly with food choice in males'') the authors go on to describe biologically why the 
  correlation would be significant for males but not females. \citep{MerilaitaandJormalainen1997}
\end{itemize}

\subsection*{Conflating statistical and substantive significance}

In almost all cases what a researcher really cares about is the size of an effect and the scientific implications of the
effect. Statistical significance does not, in and of itself, tell us anything about substantive significance. As 
\citet{Bernardietal.2017} state, ``...sociological and social significance are sacrificed on the altar of statistical 
significance''. Included in this is an absence of reported effect sizes and confidence intervals.

\begin{itemize}
  \item Lack of clarity differentiating between biological and statistical significance: {Simonetal.2017}
  \item Complete lack of effect sizes: \citep{Juriadoetal.2017}
\end{itemize}

\begin{itemize}
  \item Topics we may want to bring up
	  \subitem Including a p~value helps to avoid type-S errors
\end{itemize}

\mk{Not sure if we want any of this and how much we want to expand on any of these topics:}

@@

We (probably will) thank the Bolker lab for helpful comments on the first draft of the manuscript.

>>>>

\textbf{Abstract}: Decades of misuse and misinterpretation of Null Hypothesis Significance Testing (NHST) has resulted in
calls to abandon its use. Alternatively, we argue that statistical hypothesis tests are useful, but are not, and should not
be referred to, as tests of significance. We propose that a sensible method to reduce misconceptions about NHST and p~values 
is to refer to the results of NHST (i.e. $p < or > 0.05$) using ``statistical clarity''. A test of the null hypothesis tells 
us whether or not we can \emph{clearly} see our effect (e.g. whether we can clearly see that a correlation is positive or 
negative, or if a response in the treatment group is greater than the response in the control group). Specifically, we argue 
that the use of ``statistical clarity'' will help researchers avoid accepting the null hypothesis and reclaim the English 
definition for ``significance'', allowing for a shift in focus to substantive significance.

@@

For nearly a century statisticians and scientists have described the shortcomings of Null Hypothesis Significance Testing
(NHST) \citep[see][]{Cohen1994}. Research articles and books written in the past two decades proposing a deemphasis of the 
p~value in favor of effect sizes and confidence intervals have been cited thousands of times \citep{Cohen1994, Goodman1999, 
Wilkinson1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}. Yet, both a reliance on NHST and the misinterpretation of 
results obtained from NHST remain high. In a 2017 meta-analysis of 356 studies published in \emph{European Sociological 
Research}, \citet{Bernardietal.2017} found that 72\% of studies contained an ambiguous use of the term ``significant'', 
49\% interpreted non-significant effects as zero effects, and 44 \% failed to report an effect size in a comprehensible 
way. The misuse and misinterpretation of NHST is so pervasive that there has been a recent push for drastically reducing its 
use \citep{SzucsandIoannidis2017} or abandoning its use altogether \citep{McShaneetal.2017}. Others have called for a more
strenuous significance threshold (e.g. $p < 0.005$: \citealt{Benjaminetal.2018}); however, this change is unlikely to reduce
the misuse and misinterpretation of p~values, or the publication bias imposed by any p~value threshold \citep{Ridleyetal.2007}.

Instead, we suggest that using the language of ``statistical clarity'' would help researchers escape the trap of accepting 
the null hypothesis, reduce multiple types of misinterpretation associated with accepting the null hypothesis, and allow for
the continued use of NHST as a method of inference (e.g. see \citealt{Abelson1997} for arguments for NHST). What a test of the
null hypothesis actually tells us is whether or not we can \emph{clearly} see our effect. For example, given $H_{0} = 0$ (e.g.
the difference in our treatments is 0) what a $p < 0.05$ tells us is whether or not we can clearly see the sign of the effect. 
Therefore, results from NHST (i.e. $p < or > 0.05$) should be described using ``statistical clarity'' instead of ``statistical 
significance''. For example, equating $p < 0.05$ with a zero effect would be much harder to make once a researcher has written 
the following sentence: ``The difference between our control and treatment group was not statistically clear ($p = 0.24$). 
Additionally, the use of larger p~values as increased evidence for the null would be more difficult if the following language
was used: ``We were unable to clearly determine how our control and sham treatments differed''. Finally, in almost all cases 
what a researcher really cares about is the size of an effect and the scientific implications of the effect.  As 
\citet{Bernardietal.2017} state, ``...sociological and social significance are sacrificed on the altar of statistical 
significance''. If ``significant'' were to reclaim its common English definition, conflating $p < 0.05$ with substantive
significance would occur less frequently.

\mk{Language here needs a lot of work. Both in terms of clarity (haha) (use of ``confidence'' may be a bit confusing) and 
because I am not sure that everything I am saying here is 100\% correct. Also not sure if this is how this section should
start. Finally, portions of this have already been said in the paragraphs above.}

A p~value is not a property of reality (e.g. a biological phenomena), but is a property of a data set and the statistical test
used. A p~value must therefore be interpreted and reported as a measure that quantifies our confidence about what we can see
clearly about reality given our data. A $p < 0.05$ tells us that we are confident in in our ability to clearly see the sign
of our effect (at an $\alpha = 0.05$ cutoff). While a direct illustration of this clarity is shown with the confidence 
interval associated with $p < 0.05$, which will not overlap 0 (\mk{In nearly all cases?}), a p~value is a useful tool to 
quantify this clarity. A p~value does not tell us about the substantive (e.g. biological) importance of our finding. Because
the p~value is a property of the data, and a true zero,``null effect'' does not exist in the vast majority of scientific 
fields \citep{Meehl1990, Tukey1991, Cohen1994}, $p < 0.05$ can be obtained for a very small effect with a large enough 
sample size. Yet, references to a ``null effect'' can still be found in high impact journals \citep[e.g.][]{Feldonetal.2017}. 
To emphasize that effects are not truly zero, descriptions of statistical results using the language of clarity should begin 
with reference to the effect. For example, ``The difference between the control and treatment group was not statistically 
clear.'' In Table~1 we provide examples from the literature of language that misinterprets p~values and a sensible way of
rewriting these statements to reduce confusion.

Another common mistake is using a larger p~value as stronger evidence for accepting the null hypothesis. A larger p value 
actually corresponds to less clarity and thus the possibility of an undetected larger effect (see Table~1 for an example). In 
2006 Gelman and Stern wrote that the difference between statistically significant and statistically non-significant is not 
itself significant. Rewritten using ``clarity'' this statement indicates that if a single result is not statistically 
clear, any comparison made between that unclear result and a clear result will not itself be a clear comparison. An example
taken from the literature is in Table~1.

\mk{My table formatting skills are pretty bad}

\begin{framed}
\begin{tabular}{p{7.5cm}p{7.5cm}}
     Language in published journal articles & Rewritten using the language of ``clarity'' \\
     	\hline
        \\
     \multicolumn{2}{c}{$\mathbf{p > 0.05 \neq}$ \textbf{zero effect}} \\
     	No relationship between individual genetic diversity and prevalence of avian malaria... & 
        The relationship between individual genetic diversity and prevalence of avian malaria was not
        statistically clear \\
        \\
     \multicolumn{2}{c}{\textbf{Using larger p~values as more evidence for something}} \\
        ...yielded nonsignificant differences on measures of those skills with P values of 
        at least 0.2, and most in the $0.7 \leq P \leq 0.9$ range. & 
        ...the differences between skills developed in treatment and control individuals
        were bot statistically clear ($p > 0.05$) \\
        \\
     \multicolumn{2}{c}{\textbf{The difference between ``clear'' and ``not clear'' is ``not clear''}} \\
        Substrate choice correlated significantly with food choice in males $(r_{s} = 0.32, N = 44, P < 0.05)$ 
        but not in females $(r_{s} = 0.32, N = 24, NS)$. ... The authors discuss the biological reasons why
        there is a relationship between these variables in males, but not in females & 
        Substrate choice was positively correlated with food choice in males $(r_{s} = 0.32, N = 44, P < 0.05)$,
        but this relationship was not statistically clear in females $(r_{s} = 0.32, N = 24, NS)$. A biological
        description comparing the correlation in males vs females is inappropriate without a direct test (in
        this case with the exact same correlation estimate, this is not a sensible test). \\
\end{tabular}
\end{framed}
   \captionof{table}{Rewriting misinformed language in the published literature using our proposed updated language.}

\subsection*{Non-inferiority tests}

\subsection*{Focal vs non-focal hypotheses}

\subsection*{Previous suggestions}

\section*{Lists}

\begin{itemize}
  \item Topics we may want to bring up
	  \subitem Including a p~value helps to avoid type-S errors
      \subitem usefulness of p --- use in splines: non-ridiculous standard to determine if we can see a pattern
      \subitem difficulty in interpreting confidence intervals?
\end{itemize}

\begin{itemize}
  \item Examples
  \subitem Equating $p > 0.05$ with no difference: \citep{Ortegoetal.2007, Singhetal.2017}
  \subitem Lack of clarity differentiating between biological and statistical significance: {Simonetal.2017}
  \subitem Complete lack of effect sizes: \citep{Juriadoetal.2017}
  \subitem Setting out to accept the null: \citep{Karulinetal.2015}
  \subitem ``...establish the \emph{a priori} selection of an effect level that is considered unacceptably toxic:'' \citep{Dentonetal.2011}
  \subitem With the \emph{correlation coefficient} for males and females (but significance for only one sex---``substrate 
  choice correlated significantly with food choice in males'') the authors go on to describe biologically why the 
  correlation would be significant for males but not females. \citep{MerilaitaandJormalainen1997}
\end{itemize}

@@

@@

We thank members of the Dushoff and Bolker labs for helpful comments on the first draft of the manuscript.
